---
title: "Data Preprocessing for Procurement Trends Module"
author: "Cathy Chu"
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

## Overview

This document outlines the data preprocessing steps for the Procurement Trends module. The process involves cleaning and transforming government procurement data from GeBIZ, with a focus on temporal patterns and trends in government spending and tender awards.

## Data Dictionary

### Temporal Variables
- `award_date`: Date of tender award (format: YYYY-MM-DD)
- `award_year`: Year of award (numeric)
- `month`: Month of award (01-12)
- `year_month`: Year and month in YYYY-MM format
- `quarter`: Quarter of the year (Q1-Q4)

### Key Metrics
- `awarded_amt`: Amount awarded in the tender
- `agency`: Government agency name
- `supplier_name`: Supplier company name

## Data Import and Exploration

### Load packages
```{r}
pacman::p_load(dplyr, lubridate, readr, knitr)
``` 

### Import Data
```{r}
GeBiz_csv <- read_csv("../data/GovernmentProcurementviaGeBIZ.csv")
```

### Dataset Overview
```{r}
glimpse(GeBiz_csv)
```

### Temporal Coverage Analysis
```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Check date range
date_range <- GeBiz_csv %>%
  mutate(award_date = as.Date(award_date, format = "%d/%m/%Y")) %>%
  summarise(
    min_date = min(award_date, na.rm = TRUE),
    max_date = max(award_date, na.rm = TRUE),
    total_days = as.numeric(max_date - min_date)
  )
kable(date_range)

# Check monthly distribution
monthly_counts <- GeBiz_csv %>%
  mutate(
    award_date = as.Date(award_date, format = "%d/%m/%Y"),
    year_month = format(award_date, "%Y-%m")
  ) %>%
  count(year_month) %>%
  arrange(year_month)
kable(head(monthly_counts, 5))
```

## Data Processing

### Data Cleaning

Before analysis, we conducted exploratory data analysis to determine data quality and temporal coverage.

Our analysis revealed that some entries have "unknown" listed as the supplier_name with zero values in awarded_amt. These represent tenders that failed to receive a successful bid. We will exclude these entries from our analysis.

During data wrangling, we will:

-   Convert dates to proper date type format
-   Filter out entries with no awarded tenders
-   Add columns for Year, Month, and Quarter to facilitate temporal analysis
-   Handle any missing values in key fields

### Data Wrangling
```{r}
#| code-fold: true
#| code-summary: "Show the code"
#filter out Unknown value (no tender awarded)
GeBiz_full <- GeBiz_csv %>%
  filter(supplier_name != "Unknown")

# Convert dates
GeBiz_full$award_date <- as.Date(GeBiz_full$award_date, , format = "%d/%m/%Y")

# Remove white space if have.
GeBiz_full$agency <- trimws(GeBiz_full$agency)
GeBiz_full$supplier_name <- trimws(GeBiz_full$supplier_name)

# Handle any missing values
GeBiz_full <- GeBiz_full[!is.na(GeBiz_full$agency) & !is.na(GeBiz_full$supplier_name), ]

# Add year and month columns for time-based filtering
GeBiz_addtime <- GeBiz_full %>%
  mutate(award_year = format(award_date, "%Y"),
         month = format(award_date, "%m"),
         year_month = format(award_date, "%Y-%m"),
         quarter = quarters(award_date))
```

```{r}
kable(head(GeBiz_addtime))
```


::::: goals
::: goals-header
Note
:::

::: goals-container
-   The data now has 17,855 rows and 11 columns. Less rows from the original dataset because the unknown entries have been removed.
-   Temporal variables have been added to facilitate trend analysis.
-   The data is now ready for temporal analysis and visualization.
:::
:::::


